General constraints on magic distillation protocols from statistical mechanics

### 2021-05-04 appendix.tex
some quick fixes

\begin{proposition}\label{thm:aproperties}
    For any dimension $d$, the phase-point operators satisfy:
    \begin{enumerate}
        \item[(i)]\label{en:a1} Hermiticity and unitarity: $A_{\bmx}^\dagger = A_{\bmx} = A_{\bmx}^{-1}$;
	    \item[(ii)]\label{en:a2} Closure under transposition: $A_{(x, p)}^T = A_{(x, p)}$; \ddd{[FIX]}
	    \item[(iii)]\label{en:a3} Unit trace for odd $d$: $\tr[A_{\bmx}] = 1$;
	    \item[(iv)]\label{en:a4} Completeness relation: $\sum_{\bmz \in \cal{P}_d} A_{\bmz} = d\id$;
	    \item[(i)]\label{en:a5} Orthogonality: $\tr[A_{\bmx}^\dagger A_{\bm{x'}}] = d \delta_{\bmx,\bm{x'}}$.
	\end{enumerate}
\end{proposition}
\begin{proof}
	All properties follow from the definition in~\cref{eq:ax} along with properties of the displacement operator $D_{\bmx}$ and can be found in the literature, e.g.~\cite{cit:veitch,Vourdas_2004,cit:gross3}
\end{proof}

@@

With the help of this lemma, we directly arrive at 
\begin{theorem}\label{thm:bounds_strict}
	Given fixed positive integer $n$ and probability $p$, $\Phi_+, \Phi_-$ satisfy the following bounds:
	\begin{align*}
		\begin{split}
		&\text{1. } \Phi_+(m; n, p) \geq \sum\limits_{\ell=0}^{m/2}\left[ 16\ell\left(1-\frac{2\ell}{n}\right) \right]^{-\frac{1}{2}} 2^{-n\ent{\frac{2\ell}{n}}{p}}, \\
		&\hspace{14pt} \text{for all even } m\in [2,n] \\
		&\text{2. } \Phi_+(m; n, p) \leq \sum\limits_{\ell=0}^{m/2}\left[ 4\pi\ell\left(1-\frac{2\ell}{n}\right) \right]^{-\frac{1}{2}} 2^{-n\ent{\frac{2\ell}{n}}{p}}, \\
		&\hspace{14pt} \text{for all even } a\in [2,n] \\
		&\text{3. } \Phi_-(m; n, p) \geq \sum\limits_{\ell=1}^{(m-1)/2}\left[ 16(\ell+1)\left(1-\frac{2\ell+1}{n}\right) \right]^{-\frac{1}{2}} \times \\
		&\hspace{14pt} \times 2^{-n\ent{\frac{2\ell+1}{n}}{p}},\ \text{for all odd }a\in [1,n] \\
		&\text{4. } \Phi_-(m; n, p) \leq \sum\limits_{\ell=1}^{(m-1)/2}\left[ 4\pi(\ell+1)\left(1-\frac{2\ell+1}{n}\right) \right]^{-\frac{1}{2}} \times \\
		&\hspace{14pt} \times 2^{-n\ent{\frac{2\ell+1}{n}}{p}},\ \text{for all odd } a\in [1,n]
		\end{split}
	\end{align*}
\end{theorem}
\begin{proof}
	All four statements follow from application of~\cref{lem:comb_bounds} on the combinatorial coefficient and the defintion of relative entropy given in~\cref{eq:ent_rel}
\end{proof}

\ddd{[Remainder bounds via integrals? This might be of use for others, if they wanted to build on this line.]}
\subsection{Lorenz curve coordinates in the unital fragment}\label{app:lcsu_coord}
The Wigner distribution of the $n$--copy qutrit maximally mixed state $\left(\id/3\right)^{\otimes n}$ is the uniform probability distribution over the phase space, consisting of $9^n$ components equal to $9^{-n}$.
The Wigner distribution of the 1-copy $\epsilon$--noisy Strange state $\rho_{\rm{S}}(\epsilon)$ in the unital fragment consists of some permutation of a single negative component
\begin{equation}
	- v(\epsilon) \coloneqq - \left( \frac{1}{3} -\frac{4}{9}\epsilon \right),
\end{equation} 
and $8$ positive components
\begin{equation}
	u(\epsilon) \coloneqq \frac{1}{6} -\frac{1}{18}\epsilon.
\end{equation}
where in the unital fragment we need the condition $0 \leq \epsilon < 3/4$, so that the state contains some Wigner negativity ($-v < 0$).
It is also clear that $v \geq u$ in the interval $0 \leq \epsilon \leq 3/7$, while $u > v$ in the interval $3/7 < \epsilon < 3/4$.

@@

We can unify the indices, by introducing a single index
\begin{equation}
    I(i,j,k) \coloneqq k + \left[ (j-1) + \sum_{\ell=0}^{i-1} m_{\ell}(n', \epsilon') \right]9^{n-n'},
\end{equation}
so that $I=1,2,\dots, 9^{n}$.
The elbow coordinates correspond to 
\begin{equation}
	I(i, m_{i}(n', \epsilon'), 9^{n-n'}) = \sum_{\ell=0}^{i} m_{\ell}(n', \epsilon'),\ i= 0,\dots,n'.
\end{equation}
The index function $I$ is bijective, i.e.
\begin{equation}
	(i,j,k) = (i',j',k') \Leftrightarrow I(i,j,k) = I(i',j',k').
\end{equation}

>>>>

\begin{proposition}\label{thm:aproperties}
    For any dimension $d$, the phase-point operators satisfy:
    \begin{enumerate}
        \item[(i)]\label{en:a1} Hermiticity and unitarity: $A_{\bmx}^\dagger = A_{\bmx} = A_{\bmx}^{-1}$;
	    \item[(ii)]\label{en:a2} Closure under transposition: $A_{(x, p)}^T = A_{(x, -p)}$;
	    \item[(iii)]\label{en:a3} Unit trace for odd $d$: $\tr[A_{\bmx}] = 1$;
	    \item[(iv)]\label{en:a4} Completeness relation: $\sum_{\bmz \in \cal{P}_d} A_{\bmz} = d\id$;
	    \item[(i)]\label{en:a5} Orthogonality: $\tr[A_{\bmx}^\dagger A_{\bm{x'}}] = d \delta_{\bmx,\bm{x'}}$.
	\end{enumerate}
\end{proposition}
\begin{proof}
	All properties follow from the definition in~\cref{eq:ax} along with properties of the displacement operator $D_{\bmx}$ and can be found in the literature, e.g.~\cite{cit:veitch,Vourdas_2004,cit:gross3}
\end{proof}

@@

With the help of this lemma, we directly arrive at 
\begin{theorem}\label{thm:bounds_strict}
	Given fixed positive integer $n$ and probability $p$, $\Phi_+, \Phi_-$ satisfy the following bounds:
	\begin{align*}
		\begin{split}
		&\text{1. } \Phi_+(m; n, p) \geq \sum\limits_{\ell=0}^{m/2}\left[ 16\ell\left(1-\frac{2\ell}{n}\right) \right]^{-\frac{1}{2}} 2^{-n\ent{\frac{2\ell}{n}}{p}}, \\
		&\hspace{14pt} \text{for all even } m\in [2,n] \\
		&\text{2. } \Phi_+(m; n, p) \leq \sum\limits_{\ell=0}^{m/2}\left[ 4\pi\ell\left(1-\frac{2\ell}{n}\right) \right]^{-\frac{1}{2}} 2^{-n\ent{\frac{2\ell}{n}}{p}}, \\
		&\hspace{14pt} \text{for all even } a\in [2,n] \\
		&\text{3. } \Phi_-(m; n, p) \geq \sum\limits_{\ell=1}^{(m-1)/2}\left[ 16(\ell+1)\left(1-\frac{2\ell+1}{n}\right) \right]^{-\frac{1}{2}} \times \\
		&\hspace{14pt} \times 2^{-n\ent{\frac{2\ell+1}{n}}{p}},\ \text{for all odd }a\in [1,n] \\
		&\text{4. } \Phi_-(m; n, p) \leq \sum\limits_{\ell=1}^{(m-1)/2}\left[ 4\pi(\ell+1)\left(1-\frac{2\ell+1}{n}\right) \right]^{-\frac{1}{2}} \times \\
		&\hspace{14pt} \times 2^{-n\ent{\frac{2\ell+1}{n}}{p}},\ \text{for all odd } a\in [1,n]
		\end{split}
	\end{align*}
\end{theorem}
\nick{Should we keep this theorem??}
\begin{proof}
	All four statements follow from application of~\cref{lem:comb_bounds} on the combinatorial coefficient and the defintion of relative entropy given in~\cref{eq:ent_rel}
\end{proof}

\subsection{Theory on bounding the core functions}
The core functions are $\Phi_{\pm}$ which is an even power cumulative function. 
We can rewrite it as
\begin{equation}
\Phi_{\pm}(m; n, a) = \frac{1}{2}( \Phi(m; n, a) \pm (1+a)^{-n} S(m; n, a) ),
\end{equation}
where we have substituted $a = p/(1-p)$. 
$\Phi$ is the standard cumulative function 
\begin{equation}
	\Phi(m; n, a) = (1+a)^{-n} \sum_{k=0}^m \binom{n}{k} a^k,
\end{equation}
and the remainder term is
\begin{equation}
S(m; n, a) \coloneqq \sum_{k=0}^m \binom{n}{k} (-a)^k.
\end{equation}
We have good asymptotic bounds on the behaviour of $\Phi$~\cite{cit:ash},
\begin{lemma}\label{lem:phil_bounds}
	Given fixed $n>0$ and $p$, $\Phi_\ell$ satisfies the following bounds:
	\begin{align*}
		\begin{split}
		&\text{1. } \Phi(m; n, p) \geq \left[ 8m\left(1-\frac{m}{n}\right) \right]^{-\frac{1}{2}} 2^{-n\ent{\frac{m}{n}}{p}},\ m\in [1,n-1] \\
		&\text{2. } \Phi(m; n, p) \leq 2^{-n\ent{\frac{m}{n}}{p}},\ m\in [0,np] \\
		&\text{3. } \Phi(m; n, p) \leq 1 - \left[ 8(m+1)\left(1-\frac{m+1}{n}\right) \right]^{-\frac{1}{2}} 2^{-n\ent{\frac{m+1}{n}}{p}}, \\
		&\hspace{14pt} m\in [0,n-2]
		\end{split}
		\\
		&\text{4. } \Phi(m; n, p) \geq 1 - 2^{-n\ent{\frac{m+1}{n}}{p}},\ m\in [np+1,n-2]
	\end{align*}
\end{lemma}
\nick{Should we keep this theorem instead?? Bounds on Phi are numerically veeery loose in general, and this theorem provides even looser bounds than the previous one}

We would like some theory that estimates the value of $S(m; n, a)$ for different parameter regimes. 
We can consider the function $f(x) = (1+x)^n$ and note that $S(m; n, a)$ is the $m$'th partial sum of this expansion at the point $x=-a$.

The truncated Maclaurin series of a general function $f(x)$ is
\begin{equation}
	f(x) = f(0) + x f'(0) + \dots \frac{x^m}{m!}f^{(m)}(0) + R_m(x)
\end{equation}
with a remainder term
\begin{align}
	R_m (x)&= \int_{0}^x dt f^{(m+1)}(t) \frac{(x-t)^r}{r!} \\
	&= \frac{x^{m+1}}{(m+1)!} f^{(m+1)}(x_*),
\end{align}
where in the second expression, $x_*$ is an implicit point that lies between $0$ and $x$ that comes from the Mean Value Theorem.

Applying this to the function $f(x) = (1+x)^n$ gives
\begin{equation}
	(1+x)^n = \sum_{k=0}^m \binom{n}{k} x^k + R_m.
\end{equation}
Evaluating at $x=-a$ gives
\begin{equation}
	S(m; n, a) = (1-a)^n - R_m(-a),
\end{equation}
where the key remainder term is given by
\begin{align}
	R_m(-a) &= \int_0^{-a} dt f^{(m+1)}(t) \frac{(-a-t)^m}{m!} \\
&= \frac{(-a)^{m+1}}{(m+1)!} f^{(m+1)}(x_*).
\end{align}
We can also compute the derivative $f^{(m+1)}(x)$ explicitly,
\begin{equation}
	f^{(m+1)}(x) = (m+1)!\binom{n}{m+1}(1+x)^{n-m-1}.
\end{equation}
Therefore, we have that
\begin{align}
	R_m(-a) &= (-1)^{m+1}(m+1)\binom{n}{m+1}\int_{-a}^0 dt (1+t)^{n-m-1}(a+t)^m \\
&= \binom{n}{m+1}(-a)^{m+1}(1+x_*)^{n-m-1},
\end{align}
where in the latter expression $x_* \in [-a,0]$. 
Note that the first integral expression can be estimated via the Cauchy-Schwarz or the Hoelder inequality. 
Therefore, we can either work with an explicit form with an unknown (but bounded) parameter $x_*$, or we can use the integral form and provide concrete estimates on it value.

A very simple estimate, based on $x_*$ lying in the interval $[-a,0]$ gives
\begin{equation}
\binom{n}{m+1}(-a)^{m+1}(1-a)^{n-m-1} \leq R_m(-a) \le  \binom{n}{m+1}(-a)^{m+1}.
\end{equation}

This in turn leads to the following bounds on $\Phi_+(a,r,n)$:
\begin{align}
\hspace{-2cm}2\Phi_+(a,r,n) &\le \Phi(a,r,n)+ (1-a)^n - \frac{(-a)^{r+1}}{r!} \binom{n}{r+1} (1-a)^{n-r-1} \nonumber \\
2\Phi_+(a,r,n) & \ge\Phi(a,r,n)+ (1-a)^n - \frac{(-a)^{r+1}}{r!} \binom{n}{r+1} .
\end{align}
\nick{Fix notation !}

\subsection{Lorenz curve coordinates in the unital fragment}\label{app:lcsu_coord}
The Wigner distribution of the $n$--copy qutrit maximally mixed state $\left(\id/3\right)^{\otimes n}$ is the uniform probability distribution over the phase space, consisting of $9^n$ components equal to $9^{-n}$.
The Wigner distribution of the 1-copy $\epsilon$--noisy Strange state $\rho_{\rm{S}}(\epsilon)$ in the unital fragment consists of some permutation of a single negative component
\begin{equation}
	- v(\epsilon) \coloneqq - \left( \frac{1}{3} -\frac{4}{9}\epsilon \right),
\end{equation} 
and $8$ positive components
\begin{equation}
	u(\epsilon) \coloneqq \frac{1}{6} -\frac{1}{18}\epsilon.
\end{equation}
where in the unital fragment we need the condition $0 \leq \epsilon < 3/4$, so that the state contains some Wigner negativity ($-v < 0$).
It is also clear that $v \geq u$ in the interval $0 \leq \epsilon \leq 3/7$, while $u > v$ in the interval $3/7 < \epsilon < 3/4$.

@@

We can unify the indices, by introducing a single index
\begin{equation}
    I(i,j,k) \coloneqq k + \left[ (j-1) + \sum_{\ell=0}^{i-1} m_{\ell}(n', \epsilon') \right]9^{n-n'},
\end{equation}
so that $I=1,2,\dots, 9^{n}$.
The elbow coordinates correspond to 
\begin{equation}
	I(i, m_{i}(n', \epsilon'), 9^{n-n'}) = \sum_{\ell=0}^{i} m_{\ell}(n', \epsilon'),\ i= 0,\dots,n'.
\end{equation}
The index function $I$ is bijective, i.e.
\begin{equation}
	(i,j,k) = (i',j',k') \text{ iff } I(i,j,k) = I(i',j',k').
\end{equation}
